Post-Disturbance Tree Community Trajectories in a Neotropical Forest

### 2018-04-17 WholePlotTrajectories.Rmd
Discussion

```{r Options, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(cache=F, warning=FALSE, tidy=TRUE, tidy.opts=list(blank=FALSE, width.cutoff=40), fig.align="center")
# Installation des packages si nécessaire et chargement
Library <- function(Packages) {
  InstallAndLoad <- function(Package) {
    if (!Package %in% installed.packages()[, 1]) {install.packages(Package, repos="https://cran.rstudio.com/")}
    require(Package, character.only = TRUE)
  }
  invisible(sapply(Packages, InstallAndLoad))
}
# Ajouter les packages nécessaires ici
Library("kableExtra")
source("Scripts/DrawingTools.R")
#source("Scripts/GithubPages.R")
```

@@

The monitoring 30 years of disturbed forest communities confirmed the limited direct impact of low disturbance on species richness as it was found for impact survey of selective logging (@Cannon1998;@Baraloto2012).
Only the most disturbed plots had not fully achieved the ongoing recovery of genus richness, although all had already reached equivalent levels as those of some control plots.
Such long time richness recovery had already been observed in several logging experiments, which also highlighted the role of random disturbance damage in the richness decrease (@deAvila2015;@Hu2018).

>>>>

```{r Options, echo=FALSE, message=FALSE, warning=F}
knitr::opts_chunk$set(cache=F, warning=F, tidy=TRUE, tidy.opts=list(blank=FALSE, width.cutoff=40), fig.align="center")
# Installation des packages si nécessaire et chargement
Library <- function(Packages) {
  InstallAndLoad <- function(Package) {
    if (!Package %in% installed.packages()[, 1]) {install.packages(Package, repos="https://cran.rstudio.com/")}
    require(Package, character.only = TRUE)
  }
  invisible(sapply(Packages, InstallAndLoad))
}
# Ajouter les packages nécessaires ici
Library("kableExtra")
source("Scripts/DrawingTools.R")
#source("Scripts/GithubPages.R")
```

@@

The monitoring 30 years of disturbed forest communities confirmed the limited direct impact of low disturbance on species richness as it was found for impact survey of selective logging [@Cannon1998;@Baraloto2012].
Only the most disturbed plots had not fully achieved the ongoing recovery of genus richness, although all had already reached equivalent levels as those of some control plots.
Such long time richness recovery had already been observed in several logging experiments, which also highlighted the role of random disturbance damage in the richness decrease [@deAvila2015;@Hu2018].

### 2018-04-17 WholePlotTrajectories.tex
Discussion

\begin{verbatim}
## Warning: package 'kableExtra' was built under R version 3.3.3
\end{verbatim}

\begin{verbatim}
## Warning: package 'shape' was built under R version 3.3.3
\end{verbatim}

@@

Trajectories of Richness, Shannon and Simpson taxonomic diversity were
examined in relation to the 1989 inventories (5 years after disturbance)
(Figure \ref{fig:Fig5}). \emph{For the three indices the OLS regression
confirmed the time-dependency of the diversity variations
(P\textless{}0.01) and a covariance analysis confirmed the significant
effect of the disturbance treatment (P\textless{}0.01).}

@@

For all undisturbed plots the Rao diversity remained comparable to the
values in 1989, the trajectories of all disturbed plots showed
significant increase positively correlated with the disturbance
intensity (Figure \ref{fig:Fig6}). Although a first increase of plots
functional diversity that was all the more important than initial
disturbance was intense, 30 after disturbance the diversity was similar
among disturbed plots and significantly higher than this of control
plots. \emph{The OLS regression confirmed the time-dependency of the
diversity variations (P\textless{}0.01) and the covariance analysis
confirmed the significant effect of the disturbance treatment
(P\textless{}0.01)}.

@@

The monitoring 30 years of disturbed forest communities confirmed the
limited direct impact of low disturbance on species richness as it was
found for impact survey of selective logging
(\citet{Cannon1998};\citet{Baraloto2010}). Only the most disturbed plots
had not fully achieved the ongoing recovery of genus richness, although
all had already reached equivalent levels as those of some control
plots. Such long time richness recovery had already been observed in
several logging experiments, which also highlighted the role of random
disturbance damage in the richness decrease
(\citet{deAvila2015};\citet{Hu2018}).

The evenness of all disturbed plots and the richness of low disturbance
plots on the other hand mainly followed an asymptotic growth with a
sharp increase until 15 years after disturbance. Such increased evenness
is explained by a much more homogeneous species distribution after
disturbance. Along with the IDH, disturbance then enhanced the growth
and survival of previously infrequent species, reorganizing the typical
high dominance structure of hyperdiverse mature forests. This increase
in taxonomic diversity was accompanied by an increased taxonomic
dissimilarity to initial state along time, translating a a taxonomic
turnover benefiting to pioneers and light demanding species, as
illustrated by the communities functional shifts towards
resource-acquisitive strategies (sharp increase in the SLA, leaf
thickness and bark thickness and decrease in wood density, leaf
toughness and maximum height)
\citep{Westoby1998, Wright2004, Reich2014}. In line with the IDH, the
post-disturbance changes in abiotic environment and competitive pressure
then well favored pioneers outcompeting other species when abiotic
resources are abundant but otherwise excluded in mature forests by
long-lived, resistant and shade tolerant species. The trajectories of
proportions among seed mass classes besides revealed the enhanced growth
of small seeded, large dispersive, species and therefore the importance
of species reproductive strategy which further supported the role of
species dispersion and demographic strategy for the post-disturbance
dynamics, as suggested by the IDH
\citep{TerSteege2001, Flores2006, Haddad2008}.

@@

The divergence from the initial state however was very variable among
plots, specifically after intense disturbance, which meant a strong
dependency of plots recovery to initial composition or to abiotic
conditions. All plots, though, followed the same unimodal trajectory
which confirmed both the resilience of forests taxonomic and functional
composition and the convergence of communities towards initial states
\citep{Hubbell1999, Molino2001, Baraloto2012a}. Functional traits are
the most direct link between biodiversity and ecosystem functioning
\citep{Diaz2005} the recovery of communities functional characteristics
would go along the recovery of ecosystems processes although taxonomic
composition remain altered \citep{Guariguata2001}. \#The comparison of
observed trajectories to a null model with random traits set shuffling
showed an oriented functional shift after disturbance which suggest
additional biotic and abiotic filters for the 10 to 15 years after
disturbance. \#This means a shift from exclusive competition favoring
competitive species before disturbance to environmental filtering
selecting acquisitive and disturbance resistant species.

@@

After disturbance, communities functional characteristics followed a
similar but delayed trajectory as this of taxonomic characteristics
which difference to initial state remained higher after 30 years. This
confirms conclusions drawn for grasslands
\citep{Tilman1997, Mouillot2011} and more recently for tropical forests
\citep{Lohbeck2015, Guariguata2001}. Assuming the ``vegetation quantity
effect'', suggesting that communities functional characteristics are
defined by dominant species rather that subtle changes in biodiversity
\citep{Grime1998}, this delay between functional and taxonomic
diversities meant the rapid recovery of dominant species while this of
rare ones would be much longer. Besides, it meant that species missing
to plots taxonomic recovery would be functionally redundant with those
already in place. The functional overlap between species, typical of the
huge biodiversity of tropical forests \citep{Bellwood2006}, is then not
trivially recovered after disturbance and as a determinant of forests
resilience this questions forests resilience to multiple disturbance
\citep{Trenbath1999, Elmqvist2003, Diaz2005}.

>>>>

@@

Trajectories of Richness, Shannon and Simpson taxonomic diversity were
examined at genus level in relation to the 1989 inventories (5 years
after disturbance) (Figure \ref{fig:Fig5}).

@@

For all undisturbed plots the Rao diversity remained comparable to the
values in 1989, the trajectories of all disturbed plots showed
significant increase positively correlated with the disturbance
intensity (Figure \ref{fig:Fig6}). Although a first increase of plots
functional diversity that was all the more important than initial
disturbance was intense, 30 after disturbance the diversity was similar
among disturbed plots and significantly higher than this of control
plots.

@@

The monitoring 30 years of disturbed forest communities confirmed the
limited direct impact of low disturbance on species richness as it was
found for impact survey of selective logging
\citep{Cannon1998, Baraloto2012}. Only the most disturbed plots had not
fully achieved the ongoing recovery of genus richness, although all had
already reached equivalent levels as those of some control plots. Such
long time richness recovery had already been observed in several logging
experiments, which also highlighted the role of random disturbance
damage in the richness decrease \citep{deAvila2015, Hu2018}.

The evenness of all disturbed plots and the richness of low disturbance
plots on the other hand mainly followed an asymptotic growth with a
sharp increase until 15 years after disturbance. Such increased evenness
is explained by a much more homogeneous species distribution after
disturbance. Communities composition after disturbance is either due to
the old, pre-disturbance survivors or to the recruited trees. The
composition of old survivors proved to mirror the initial community
\citep{Herault2018}, the observed composition turnover would stem from
an enhanced the growth and survival of previously infrequent species
which, along with the IDH, reorganizes the typical high dominance
structure of hyperdiverse mature forests. This increase in taxonomic
diversity was accompanied by an increased taxonomic dissimilarity to
initial state along time, translating a a taxonomic turnover benefiting
to pioneers and light demanding species, as illustrated by the
communities functional shifts towards resource-acquisitive strategies
(sharp increase in the SLA, leaf thickness and bark thickness and
decrease in wood density, leaf toughness and maximum height)
\citep{Westoby1998, Wright2004, Reich2014}. Post-disturbance changes in
abiotic environment and competitive pressure then well favored pioneers
outcompeting other species when abiotic resources are abundant but
otherwise excluded in mature forests by long-lived, resistant and shade
tolerant species. The trajectories of proportions among seed mass
classes besides revealed the enhanced growth of small seeded, large
dispersive, species and therefore the importance of species reproductive
strategy which further supported the role of species dispersion and
demographic strategy for the post-disturbance dynamics, as suggested by
the IDH \citep{TerSteege2001, Flores2006, Haddad2008}.

@@

Plots functional diversity had recovered initial values for almost all
plots 30 years after disturbance. It followed a hump-back trajectory
that was mirror by all leaf, stem and life history traits, to the
exception of leaf chlorophyll content. This recovery of communities
functional characteristics would go along the recovery of ecosystems
processes \citep{Guariguata2001}, as functional traits are the most
direct link between biodiversity and ecosystem functioning
\citep{Diaz2005}. Similar hump-back trajectories were followed by
communities taxonomic and functional composition which argued for the
convergence of communities composition maintaining the initial
difference after the recovery cycle
\citep{Hubbell1999, Molino2001, Baraloto2012a}. Besides, desipte a
homogeneous trend among plots, the trajectories proved quite variable
among plots, specifically after intense disturbance. This meant a strong
dependency of plots recovery time and trajectory to initial composition
or abiotic conditions.

\section{The comparison of observed trajectories to a null model with
random traits set shuffling showed an oriented functional shift after
disturbance which suggest additional biotic and abiotic filters for the
10 to 15 years after
disturbance.}\label{the-comparison-of-observed-trajectories-to-a-null-model-with-random-traits-set-shuffling-showed-an-oriented-functional-shift-after-disturbance-which-suggest-additional-biotic-and-abiotic-filters-for-the-10-to-15-years-after-disturbance.}

\section{This means a shift from exclusive competition favoring
competitive species before disturbance to environmental filtering
selecting acquisitive and disturbance resistant
species.}\label{this-means-a-shift-from-exclusive-competition-favoring-competitive-species-before-disturbance-to-environmental-filtering-selecting-acquisitive-and-disturbance-resistant-species.}

@@

Composition trajectories, though, translated marked differences in
communities composition as the functional diversity had already
recovered. Such delay between functional and taxonomic dynamics, already
observed for grasslands \citep{Tilman1997, Mouillot2011} and more
recently for tropical forests \citep{Lohbeck2015, Guariguata2001}.
Assuming the ``vegetation quantity effect'', suggesting that communities
functional characteristics are defined by dominant species rather that
subtle changes in biodiversity \citep{Grime1998}, this delay meant a
faster recovery of dominant species while this of infrequent ones would
be much longer. These infrequent species missing to plots taxonomic
recovery would consequently be functionally redundant with those already
in place. The functional overlap between species, typical of the huge
biodiversity of tropical forests \citep{Bellwood2006}, is then not
trivially recovered after disturbance and as a determinant of forests
resilience this questions forests resilience to multiple disturbance
\citep{Trenbath1999, Elmqvist2003, Diaz2005}.
